\chapter{The present tense of regular  verbs, irregular verbs, and  verbs with spelling changes}

    \section{}

    \begin{itemize}
        \item to conjugate a verb in any tense you need two pieces of information: the root of the verb and the ending (suffix) that contains the information about the action framed in a form the speaker desires to convey.
        \item the root of a word is also called the radical. In english infinitives tend to be precided by 'to', as in 'to learn', 'to say', etc. However in Spanish infinitives are instead indicated by suffixes: -ar, -er, or -ir. Thus for hablar, beber, and decidir the roots are habl-, beb-, and decid-, and the suffixes are respectively -ar, -er, and -ir.
        \item the conjugations are numbered: -ar is called the first conjugation, -er is called the second conjugation, and -ir is called the third conjugation.
        \item Thus conjugations are formed by removing the suffix from the infinitive and replacing it with the required suffix.
        \item Since conjugation contains information about the subject, subject pronouns tend to be omitted from a sentence. An example is "vendo", meaning "I sell" which states that the individual is speaking about themselves in first person, *I* = "-o" in this case.
        \item Conjugations contain information about the individual being described, the tense and the gender.
        \item Moods express the attitude of the speaker.
        \item Moods include indicative, subjunctive, conditional and imperative.
        \item The indicative mood is spoken as "modo indicativo" and is used when verbs describe situations which are real, true, or taken as fact.
        \item The suffix of the verb contains information about the following: who does the action, when it happened, and the mood of the speaker.
    \end{itemize}

    \section{Regular verbs in the present tense}

    \begin{itemize}
        \item the present tense of a verb is formed by dropping the suffix and selecting a new suffix that indicates the subject.
        \item The author recommends using the verbs "comprar", "aprender" and "recibir" as models.
        \item The author presents conjugation tables of two columns: singular and plural, iterating down the subjects: first, second and third. Thus all verbs will have 6 conjugations in the present tense.
        \item subject pronouns and their translations are as follows: yo: I, tú: you, el: he, ella: she, usted: you (formal), nosotros: we, nosotras we (fem.), vosotros: you, vosotras you, (fem.), ustedes: you (formal)
        \item some notes bout present tense conjugation: 1. yo uses the same suffix for all conjugations (first, second, third), 2. for the second and third conjugation, all third person singular forms have the same suffix 3. in spanish, subject pronouns tend to be ommited.
        \item common first conjugation verbs include: bajar, cocinar, comprar, conversar, cooperar, dibujar, escuchar, limipar, preparar, sacar, sumar, trabajar
        \item common second conjugatoin verbs include: beber, comer, comprender, correr, creer, leer, responder, romper, temer, vender
        \item common third conjugation verbs include: abrir, compartir, cubrir, decidir, discutir, escribir, repartir, subir, sufrir, vivir
    
    \section{Exercises}

    % todo
    \end{itemize}
